---
title: "Activity: Avocado Prices"
output: rmdformats::readthedown
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=12, fig.height=8, fig.path='Figs/',
                      echo=FALSE, warning=FALSE, message=FALSE,eval=TRUE, include = FALSE)
```


# Instructions


I advise you to focus particularly on:

* Setting chunk options carefully.

* Making sure you don't print out more output than you need to.

* Making your code readable and nicely formatted. 

* Thinking through your desired result **before** writing any code.

# Introduction

In this lab we're going to be looking at avocado prices. 

The dataset comes to us from kaggle and represents weekly retail scan data.  It is available to download from Canvas under the name `avocado.csv`.  A description of the data can be found [here.](https://www.kaggle.com/neuromusic/avocado-prices)

```{r, echo = FALSE}
library(tidyverse)
avo <- read_csv("https://www.dropbox.com/s/vsc1dkosz6nwake/avocado.csv?dl=1")
```

0) Create an R Markdown file, load the data, and declare your package dependencies. 

1) Briefly describe the dataset.  What information does it contain?

2) Clean the data in any way you see fit. In particular, look carefully at the `region` variable.  Notice that the category `LosAngeles` is contained in `California`, which is contained in `West`, which is contained in `TotalUS`.  Think about how you want to handle this issue in your analysis.



# Exercises


1) Which major region sold the most organic small Hass avocados in 2017? 

2) Use `separate()` to split the `Date` variable into year, month, and day.  In which month is the highest volume of avocado sales?

3) Which metro area regions sold the most total avocados?  Plot side-by-side boxplots of Total Volume for only the 5 regions with the highest averages for the Total Volume variable. 


# Reshaping

The following four California regions are in this dataset:  LosAngeles, SanDiego, Sacramento, SanFrancisco.  Answer the following questions about the California regions only.  In which regions is the price of organic versus conventional avocados most different?  Support your answer with a few summary numbers and a plot.  

*Hint: This questions will require restructuring of your data!*




