---
title: "Hypothesis Testing"

output:
  xaringan::moon_reader:
    lib_dir: libs
    css: xaringan-themer.css
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE, fig.height = 6, fig.width = 8)
options(htmltools.dir.version = FALSE)
library(tidyverse)
library(cowplot)
```

class: center, middle

# Overview
## R: Hypotheses
## E: Significance Level
## C: Test Statistic
## A: P-value
## P: Conclusion
---
## Research Question -> Hypotheses

```{r, echo = FALSE}
mu1 = 400
mu2 = 400

sigma1 = runif(1, 100, 200)
n1 = sample(20:50, 1)

sigma2 = runif(1, 100, 200)
n2 = sample(20:50, 1)

sample1 = rnorm(n1, mu1, sigma1)
sample2 = rnorm(n2, mu2, sigma2)

grizzlies <- data.frame(cbind(c(sample1, sample2), c(rep("Californian", n1), rep("Mexican", n2))))
names(grizzlies) = c("Weight", "Type")

grizzlies <- grizzlies %>% na.omit() %>% mutate(
  Weight = as.numeric(Weight),
  Type = factor(Type)
)

```



**RQ:** Do California grizzlies have a different average weight than Mexican grizzlies?

--

* What if the answer is *no*?

    $$H_0: \mu_C = \mu_M$$
    "Californian and Mexican grizzlies have the same true mean weight."

--

* What if the answer is *yes*?

$$H_A: \mu_C \neq \mu_M$$

"Californian and Mexican grizzlies do not have the same true mean weight."

---
## Evidence -> Significance level

We plan to take a sample of CA and MX grizzlies.

* What data would support the **alternate hypothesis**?

--

If the average weight of the CA grizzlies were much larger than the average weight of MX grizzlies

*or*

If the average weight of the MX grizzlies were much larger than the average weight of CA grizzlies

-- 

* How strong would the evidence need to be to **convince you**?

$$\alpha = 0.05$$

"We perform this test at the 0.05 significance level."

"We will reject the null if the p-value is below 0.05."

"We need to see data that has a 5% or less chance of happening, under the null, to be convinced."

---
## Collect Data -> Test Statistic

```{r, echo = TRUE, include = TRUE, message = FALSE, warning = FALSE}
set.seed(1847)
grizzlies %>% group_by(Type) %>% summarize_at(vars(Weight), funs(mean, sd, n()))
```

--

$$\bar{x}_C = 40.1, \bar{x}_M = 45.1$$
$$s_C = 21.4, s_M = 27.3$$
$$n_C = 44, n_M = 40$$

$$SE_{diff} = `r sqrt((21.4)^2/44 + 27.3^2/40)`$$

---
## Test Statistic

The **test statistic** is the one value that summarize all of our evidence.

* t-score or t-statistic

* z-score or z-statistic

* Chi-sq statistic

* ... many more!

--

We have to use a test statistic that lets us find a **p-value**!

That means we have to know the **distribution** of the test statistic, for it to be useful!

* t/z-score:  The CLT helps us.

* Comparing across categories:  Permutation to simulate distribution

* Other statistics:  *mathemagic* + computers

---
## Analysis -> p-value

What are the chances of getting this much evidence, or more, just by luck?  (If the null is true.)

$$t = \frac{40.1 - 45.1}{5.33} = -0.94$$

* What t-score did we *expect* under the null?

* What is the standard deviation of the *t-score* curve?

* What area would be *more evidence* for the **alternate hypothesis**?

---
class: center, middle

# Minor detail: Two-sided or one-sided?

---
## Two-sided vs one-sided

What if your research question had been...

> Do California grizzlies weigh less on average than Mexican grizzlies?

... or ... 

> Do California grizzlies weigh more on average than Mexican grizzlies?

---
## Two-sided vs one-sided

What if your research question had been...

> Do California grizzlies weigh less on average than Mexican grizzlies?

$$H_a:  \mu_C < \mu_M$$

p-value:  0.16

--

> Do California grizzlies weigh less on average than Mexican grizzlies?

$$H_a:  \mu_C > \mu_M$$

p-value:  0.84

---
## Analysis -> p-value

What are the chances of getting this much evidence, or more, just by luck?  (If the null is true.)

$$t = \frac{40.1 - 45.1}{5.33} = -0.94$$

*Estimated* p-value:  0.32

*Exact* p-value: 

```{r}
pt(-0.94, df = 39)*2
```

--

"There is a 35% chance of seeing a difference of sample mean weights of 5lbs or greater, if in fact there is no true difference in mean weights."

---
## Presentation/Conclusion

"With a t-score of -0.94 and 39 degrees of freedom, we get a p-value of 0.35.  We fail to reject the null at the 0.05 level.  We conclude that there is no evidence of a true difference in mean weights between Californian and Mexican grizzlies."

--

## THESE ARE BAD:

* "There is 35% chance that the null is true"

* "We found evidence that there is no difference in mean weights."

* "After seeing the p-value, we decided to use a signficance level of 0.10 instead."

